---
title: "Clustering_results_report"
# output: html_document
# date: '2024-01-24'
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    # theme: united
    highlight: tango
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
---


```{r, echo=FALSE, results='hide'}
source("../../R_code/Rfunctions.R")
```

```{r, include=FALSE}
library(tidyverse)
```

# Clustring Processed data without auxiliary variables


```{r loadData, include=FALSE}
miceObj <- readRDS("../SavedObjects/miceObj.rds")
dataList <- miceRanger::completeData(miceObj)
my_imp_data <- dataList[[1]]
# select only asthamtic data
my_imp_data <- my_imp_data %>% filter(asthma == 'Yes')
```


## Imporved Deep Empedding Clustering on the encoded data 5 clusters {.tabset}

```{r OH_IDEC_6,  echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# get the clusters
result_OH_IDEC_No_aux_6 <- read.csv("../SavedObjects/result_as_HPFcIDEC_clusters_5.csv")
# IDEC results on OH
results_cluster_data <- get_ClusterResults(OriginalData = my_imp_data, Labels = result_OH_IDEC_No_aux_6$cluster) 
# Remove asthma variable from the results
results_cluster_data <- results_cluster_data %>% select(-asthma)
cat(
  knitr::knit_child('Clustering_res_temp.Rmd', envir = environment(), quiet = TRUE), sep = '\n'
  )
```

## Imporved Deep Empedding Clustering on the encoded data 4 clusters {.tabset}

```{r OH_IDEC_4,  echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# get the clusters
result_OH_IDEC_No_aux_4 <- read.csv("../SavedObjects/result_as_HPFcIDEC_clusters_4.csv")
# IDEC results on OH
results_cluster_data <- get_ClusterResults(OriginalData = my_imp_data, Labels = result_OH_IDEC_No_aux_4$cluster) 
# Remove asthma variable from the results
results_cluster_data <- results_cluster_data %>% select(-asthma)
cat(
  knitr::knit_child('Clustering_res_temp.Rmd', envir = environment(), quiet = TRUE), sep = '\n'
  )
```

## Imporved Deep Empedding Clustering on the asthmatic encoded data 3 clusters {.tabset}

```{r OH_IDEC_3,  echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# get the clusters
result_OH_IDEC_No_aux_3 <- read.csv("../SavedObjects/result_as_FcIDEC_clusters_3.csv")
# IDEC results on OH
results_cluster_data <- get_ClusterResults(OriginalData = my_imp_data, Labels = result_OH_IDEC_No_aux_3$cluster) 
# Remove asthma variable from the results
results_cluster_data <- results_cluster_data %>% select(-asthma)
cat(
  knitr::knit_child('Clustering_res_temp.Rmd', envir = environment(), quiet = TRUE), sep = '\n'
  )
```


## Deep Empedding Clustering on the encoded data 5 clusters {.tabset}

```{r OH_DEC_6,  echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# get the clusters
result_OH_DEC_No_aux_6 <- read.csv("../SavedObjects/result_as_HPDEC_clusters_5.csv")
# IDEC results on OH
results_cluster_data <- get_ClusterResults(OriginalData = my_imp_data, Labels = result_OH_DEC_No_aux_6$cluster) 
# Remove asthma variable from the results
results_cluster_data <- results_cluster_data %>% select(-asthma)
cat(
  knitr::knit_child('Clustering_res_temp.Rmd', envir = environment(), quiet = TRUE), sep = '\n'
  )
```

## Deep Empedding Clustering on the encoded data 4 clusters {.tabset}

```{r OH_DEC_4,  echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# get the clusters
result_OH_DEC_No_aux_4 <- read.csv("../SavedObjects/result_as_HPDEC_clusters_4.csv")
# IDEC results on OH
results_cluster_data <- get_ClusterResults(OriginalData = my_imp_data, Labels = result_OH_DEC_No_aux_4$cluster) 
# Remove asthma variable from the results
results_cluster_data <- results_cluster_data %>% select(-asthma)
cat(
  knitr::knit_child('Clustering_res_temp.Rmd', envir = environment(), quiet = TRUE), sep = '\n'
  )
```
## Deep Empedding Clustering on the asthmatic encoded data 3 clusters {.tabset}

```{r OH_DEC_3,  echo=FALSE, results='asis', message=FALSE, warning=FALSE}
# get the clusters
result_OH_DEC_No_aux_3 <- read.csv("../SavedObjects/result_as_HPDEC_clusters_3.csv")
# IDEC results on OH
results_cluster_data <- get_ClusterResults(OriginalData = my_imp_data, Labels = result_OH_DEC_No_aux_3$cluster) 
# Remove asthma variable from the results
results_cluster_data <- results_cluster_data %>% select(-asthma)
cat(
  knitr::knit_child('Clustering_res_temp.Rmd', envir = environment(), quiet = TRUE), sep = '\n'
  )
```



